---
title: "Displacement Locations UNHCR Data"
output:
  md_document:
    variant: markdown_github
always_allow_html: true
subtitle: Data from Zhou and Shaver APSR 2021
author: ' '
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r setup, echo = FALSE}
knitr::opts_chunk$set(tidy = TRUE, 
               echo = FALSE, 
               warning = FALSE, 
               message = FALSE)
```

```{r load data}

## packages
require(knitr)
require(tidyverse) 
require(patchwork)
require(sf) 
require(rgdal)
require(maps)
require(mapdata)
require(maptools)
require(ggmap)

##### Load displacement locations data #####
allsites <- read.csv("displacement_locations_unhcr.csv", header = TRUE) 

## ggplot theme
yy_theme <- function(){
  theme(panel.background = element_blank(),
          legend.title = element_blank(), 
          #plot.title = element_text(size = 10),
          panel.border = element_rect(colour = "gray50", fill=NA, size=.11),
          legend.position = 'none')
}

cbPalette <- c("#E69F00", "#56B4E9", "#009E73", "#0072B2", "#D55E00", "#CC79A7")
```

# Dataset of Displacement Locations (UNHCR)

This repository contains the data of displacement (IDP, refugee, returnee, asylum seeker) locations as featured in:

Zhou, Yang-Yang and Andrew Shaver. (2021). [Reexamining the Effect of Refugees on Civil Conflict: A Global Subnational Analysis.](https://doi.org/10.1017/S0003055421000502) American Political Science Review, 1-22. doi:10.1017/S0003055421000502

If using this dataset, please cite this paper and acknowledge the UNHCR Field Information and Coordination Support Section (FICSS). 

## Overview

**displacement_locations_unhcr.csv** is a global dataset from the UNHCR Field Information and Coordination Support Section (FICSS). This dataset encompasses the universe of refugee, returnee, IDP, and asylum seeker locations that the UNHCR is aware of (even if they are not actively working in these locations), and it includes historic and currently open sites. The earliest recorded locations are from 1966, and the data was last updated by FICSS on January 6, 2020. 

To construct these data, UNHCR country teams and field staff report the location of any known displaced communities to FICSS.

With the help of a team of research assistants, we manually validated and corrected refugee and IDP locations that were open during our study period from 1990--2018. Outside of this time span, we have not made corrections.

Thus, we share this dataset on Github not only to provide researchers with global data on historic and current displacement locations, but also to make this a "living dataset," where experts can send us corrections and where the most updated version is always publicly available. 


## Descriptives

**displacement_locations_unhcr.csv** has `r nrow(allsites)` observations. 

The table and map below show the displacement locations by migration type and whether they are open or closed as of late 2019/early 2020.

```{r tab_type, eval=TRUE, results='asis'}

num_total <- allsites %>%
  group_by(type) %>%
  count(name = "N") %>%
  drop_na()
num_status <- allsites %>%
  group_by(type, status) %>%
  count() %>%
  drop_na() %>%
  mutate(status = case_when(status == 0~"Closed", status == 1~"Open")) %>%
  pivot_wider(names_from = status, values_from = n, values_fill = 0)

left_join(num_total, num_status) %>%
  rename(Type = type) %>%
  kable(caption="Number of locations by type", "pipe") 

```

```{r map_type, eval=TRUE, fig.width = 10, fig.height = 6}

allsites$status <- as.factor(allsites$status)

allsites.shp <- st_as_sf(x = allsites %>% drop_na(POINT_X, POINT_Y), 
                        coords = c("POINT_X", "POINT_Y"),
                        crs = "+proj=longlat +ellps=WGS84 +datum=WGS84")

WorldData <- map_data('world') %>% filter(region != "Antarctica") %>% fortify

ggplot() +
  geom_map(data = WorldData, map = WorldData,
                  aes(x = long, y = lat, group = group, map_id=region),
                  fill = "white", colour = "gray20", size=0.5) +
  geom_sf(data = allsites.shp, aes(colour = type)) +
  theme_classic() + 
  scale_colour_manual(values = cbPalette) + 
  labs(colour = "Location Type") + 
  theme(legend.position = "bottom")

ggplot() +
  geom_map(data = WorldData, map = WorldData,
                  aes(x = long, y = lat, group = group, map_id=region),
                  fill = "white", colour = "gray20", size=0.5) +
  geom_sf(data = allsites.shp %>% mutate(status = case_when(status == 0~"Closed", status == 1~"Open")),
          aes(colour = status)) +
  theme_classic() + 
  scale_colour_manual(values = c("#0072B2", "#D55E00")) + 
  labs(colour = "Status") + 
  theme(legend.position = "bottom")

```

## Codebook

* *pcode*: location ID
* *Country*: country of the location
* *name*: name of the location
* *name_alt*: alternative name of the location if one exists
* *loc_type*: type of location -- Asylum Seeker Accommodation, Asylum Seeker Centre, Asylum Seeker Location, Asylum Seeker Settlement, IDP Accommodation, IDP Camp, IDP Center, IDP Dispersed, IDP Location, IDP Settlement, IDP Urban location, Refugee Accommodation, Refugee Camp, Refugee Center, Refugee Dispersed, Refugee Location, Refugee Settlement, Refugee Urban location, Returnee Accommodation, Returnee Camp, Returnee Center, Returnee Location, Returnee Settlement, Unknown. Most types are self explanatory, but according to our discussions with UNHCR staff, whether a location is designed as a "camp" and "settlement" is generally due to host country government preferences.
* *loc_subtype*: some locations have additional designation types -- Aliens Police Detention Facility, Asylum Integration Center  (only for recognized refugees), Border Guard/Police Detention Facility, CHT (Centres d'Hebergement Temporaire / temporary reception centre), Home for Separated Children Seeking Asylum, Open reception facility for asylum-seekers, Open Shelters for Persons Authorized to Stay (Subsidiary form of protection).
* *createdate*: date the location was created. Many dates are 1/1/Year, which suggest the exact date is unknown but it was within that year. 
* *closedate*: if the location is closed, the date of closure
* *updatedate*: when this data entry was updated, typically by UNHCR FICSS staff
* *source*: source of the data
* *unhcr_assist*: whether UNHCR has provided assistance to the location
* *status*: 0 if closed, 1 if open as of the updatedate
* *POINT_X*: longitude 
* *POINT_Y*: latitude 
* *adjusted_vars*: if anyone has made additional adjustments to the data from the UNHCR. E.g. observations that we (Zhou and Shaver) have made changes to.
* *type*: Asylum seeker, IDP, Returnee, Refugee, or Other

## Submit a correction to the data

If you would like to submit a correction to the data, go to Issues > New Issue > Data Entry and submit a report to help us improve the dataset. Please make sure to send us a source (e.g. a report from the UNHCR, IOM, etc.) so that we can validate the correction and for transparency. 

## Coming Soon

We will continue to publicly release related data when they are ready. Here's what we're working on now:

1. **Refugee Location-Year Population data**: for each refugee location (eg. refugee camp, settlement), this dataset will also include total population by year along with population disaggregated by gender and origin country. 

2. **Uganda refugee settlement shapefiles and population data**: for Uganda, shapefiles of refugee settlements over time, along with their total population figures for election years, along with population disaggregated by origin country. 

## Acknowledgements 

We are indebted to experts from the United Nations High Commissioner for Refugees (UNHCR): Jason Hepps, Ewen Macleod, Raouf Mazou, Brett Moore, Petra Nahmias, Htun Zaw Oo, Yvon Orand, Kimberly Roberson, Edgar Scrase, Karl Steinacker, Tayyar Sukru Cansizoglu, Vicky Tennant, Alex Tyler, and staff members of the Field Information and Coordination Support Section (FICSS) for providing data and guidance.

